---
title: "HC2 Validation: Coverage Under Heteroskedasticity"
author: "Quantitative Methods"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
---

# Purpose

This notebook is **Part 2 – Apply HC2 to heteroskedastic data and develop an N-invariant severity measure**.

Part 1 established a baseline: how the inferential score behaves when data is perfectly specified (homoskedastic). Now we ask: what happens when heteroskedasticity violates that assumption? The workflow:

1. Use the HC2 estimator selected in Part 1
2. Simulate regression data with *intentional* heteroskedasticity at varying strengths
3. Demonstrate that the raw inferential score scales with sample size (it's a test statistic, not a severity measure)
4. Develop an N-invariant alternative: the **reliability score** that measures heteroskedasticity severity independent of sample size
5. Validate the reliability score's N-invariance properties
6. Analyze coverage degradation under heteroskedasticity

---

# Section 1: Setup and Load Notebook 1 Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Load Support Functions and Configuration

```{r source-files, echo=FALSE}
# Source modular R support functions
# R/00_config.R handles package loading, parallel setup, and global constants
source("R/00_config.R", local = TRUE)              
source("R/10_dgp_and_fits.R", local = TRUE)        # Data generation and fitting
source("R/20_metrics.R", local = TRUE)              # Metric computations
source("R/40_hetero_sims.R", local = TRUE)          # Heteroskedastic simulation grid
source("R/50_invariance_and_lookup.R", local = TRUE) # N-invariance analysis utilities
source("R/60_tables_and_plots.R", local = TRUE)     # Presentation utilities
```

```{r setup-paths, echo=FALSE}
BASE_PATH <- getwd()

# Note: Global seed is set in R/00_config.R and will be used for all simulations
# The seed is automatically propagated to parallel workers via furrr_options(seed = TRUE)

# Create results directory if necessary
if (!dir.exists("results")) {
  dir.create("results", recursive = TRUE)
}
```

## Load Results from Null Calibration

```{r load-null-results, results='hide'}
# Retrieve HC type selected in null calibration
best_hc <- readRDS("results/best_hc.rds")
scaling_results <- readRDS("results/scaling_results.rds")

cat(sprintf("Selected HC estimator: %s\n", best_hc))
cat(sprintf("Stability range at optimal scaling: %.6f\n\n", 
            scaling_results[hc_type == best_hc, stability_range_inf]))
```

Output: reports the selected HC estimator and its stability range from Part 1.

---

# Section 2: The Heteroskedastic Test Case

We now introduce heteroskedasticity intentionally in the data generating process. The regression model is:
$$Y_i = 1 + \beta_x \cdot X_i + \varepsilon_i$$
where the error variance depends on $X_i$:
$$\text{Var}(\varepsilon_i) = \sigma_0^2 + \lambda \cdot X_i^2$$

Parameters we vary:
- **$\beta_x$**: effect size (strength of $X$ in predicting $Y$)
- **$\sigma_0$**: baseline noise (homoskedastic component)
- **$\lambda$**: heteroskedasticity strength (controls how variance depends on $X$)
- **$N$**: sample size

We will test all combinations to see how HC2 handles different severities of misspecification.

```{r hetero-config, results='hide'}
# Simulation grid: sample sizes, heteroskedasticity strengths, and replicates per cell
# Note: These are defined in R/00_config.R and loaded automatically.
# We display them here for confirmation.

total_hetero <- length(N_GRID_HETERO) * 
                length(HETERO_STRENGTH_GRID) * 
                length(BETA_X_GRID) * 
                length(SIGMA0_GRID) * 
                N_SIMS_PER_CELL_HETERO

cat(sprintf("Simulation grid configuration:\n"))
cat(sprintf("  Sample sizes (N):           %s\n", paste(N_GRID_HETERO, collapse = ", ")))
cat(sprintf("  Hetero strength (λ):        %d levels from %.1f to %.1f\n", 
            length(HETERO_STRENGTH_GRID), min(HETERO_STRENGTH_GRID), max(HETERO_STRENGTH_GRID)))
cat(sprintf("  Beta X values:              %s\n", paste(BETA_X_GRID, collapse = ", ")))
cat(sprintf("  Sigma0 values:              %s\n", paste(SIGMA0_GRID, collapse = ", ")))
cat(sprintf("  Replicates per cell:        %d\n", N_SIMS_PER_CELL_HETERO))
cat(sprintf("  Total simulations:          %d\n", total_hetero))
cat(sprintf("  HC estimator:               %s\n\n", best_hc))
```

Output: prints the full simulation grid configuration for transparency.

---

# Section 3: Heteroskedastic Simulation and Metrics

Execute the full heteroskedastic simulation grid. For each cell (N, λ, β, σ0), generate replicates, compute 95% confidence intervals using both classical and HC standard errors, and record coverage outcomes.

The simulation engine `run_hetero_sim_grid` (in `R/40_hetero_sims.R`) computes:
- `sr_inf`: Inferential Score $S_{\mathrm{Inf}} = \text{SE}_{\text{robust}}/\text{SE}_{\text{classic}} - 1$
- `sr_inf_adj`: Scaled Inferential Score $S_{\mathrm{Inf}}^{*} = \sqrt{N} \cdot S_{\mathrm{Inf}}$
- `coverage`: Classical CI coverage (using $\text{SE}_{\text{classic}}$)

**Note:** Ratio-based metrics ($S_{\mathrm{Rel}}$) are computed post-hoc in Section 5 after establishing the N-dependence problem.

```{r run-hetero-sims}
# Note: Random seed already set in R/00_config.R and propagated to parallel workers
# via furrr_options(seed = TRUE) for full reproducibility

hetero_sim_results <- run_hetero_sim_grid(
  N_grid = N_GRID_HETERO,
  hetero_strength_grid = HETERO_STRENGTH_GRID,
  beta_x_grid = BETA_X_GRID,
  sigma0_grid = SIGMA0_GRID,
  n_sims_per_cell = N_SIMS_PER_CELL_HETERO,
  hc_type = best_hc,
  verbose = TRUE
)

# Save raw simulation results
save_rds(hetero_sim_results, "hetero_sim_results.rds")
```

## Summary of Simulation Results

We summarize the heteroskedastic simulation results, focusing on the behavior of the **Inferential Score** $S_{\mathrm{Inf}}$ across the experimental grid before proceeding to formal analysis.

```{r simulation-summary, results='hide'}
# Basic summary statistics
cat("Simulation Results Summary\n")
cat("=" |> rep(60) |> paste(collapse = ""), "\n\n")

cat(sprintf("Total simulations:        %d\n", nrow(hetero_sim_results)))
cat(sprintf("Sample sizes (N):         %s\n", paste(sort(unique(hetero_sim_results$N)), collapse = ", ")))
cat(sprintf("Hetero strengths (λ):     %d levels\n", length(unique(hetero_sim_results$hetero_strength))))
cat(sprintf("Effect sizes (β_x):       %s\n", paste(unique(hetero_sim_results$beta_x), collapse = ", ")))
cat(sprintf("Baseline noise (σ0):      %s\n", paste(unique(hetero_sim_results$sigma0), collapse = ", ")))

# Summary of Inferential Score
cat("\n\nInferential Score (sr_inf) Summary:\n")
cat("-" |> rep(40) |> paste(collapse = ""), "\n")
print(summary(hetero_sim_results$sr_inf))

# Summary of Scaled Inferential Score
cat("\n\nScaled Inferential Score (sr_inf_adj = sqrt(N) * sr_inf) Summary:\n")
cat("-" |> rep(40) |> paste(collapse = ""), "\n")
print(summary(hetero_sim_results$sr_inf_adj))
```

Output: prints grid coverage and score summaries; the key patterns are discussed below.

### Distribution of the Inferential Score

We examine the distribution of $S_{\mathrm{Inf}}$ across all simulations and by heteroskedasticity level.

```{r sinf-distribution, fig.width=10, fig.height=4}
# Use centralized plotting function from R/60_tables_and_plots.R
plot_sinf_distribution(hetero_sim_results, lambda_subset = c(0, 0.5, 1.0, 1.5, 2.0))
```

**Key Observations:**
- Under the null ($\lambda = 0$), $S_{\mathrm{Inf}}$ is centered near zero.
- As heteroskedasticity increases, $S_{\mathrm{Inf}}$ shifts positive (robust SE > classic SE).
- The distribution widens with increasing $\lambda$, reflecting greater variability under misspecification.

# Section 4: The Problem: N-Dependence of the Inferential Score

We now investigate whether the inferential score works as a severity measure or primarily as a test statistic. The scaled version:
$$T_{S_{\mathrm{Inf}}} = \sqrt{N} \cdot S_{\mathrm{Inf}}$$
is approximately standard normal under homoskedasticity, so we can reject the null (homoskedasticity) if $T_{S_{\mathrm{Inf}}} > 1.645$.

## 4.1 Power Analysis

```{r power-analysis}
# Compute T_SInf (which is sr_inf_adj in our nomenclature)
hetero_sim_results[, T_SInf := sr_inf_adj]

# Define rejection indicator
hetero_sim_results[, reject_null := as.integer(T_SInf > 1.645)]

# Aggregate rejection rates by N and hetero_strength (averaging over beta_x and sigma0)
power_summary <- hetero_sim_results[, .(
  rejection_rate = mean(reject_null)
), by = .(N, hetero_strength)]

# Use centralized plotting function from R/60_tables_and_plots.R
print(plot_power_curves(power_summary))
```

**Interpretation:**
- **Null behavior**: When data is homoskedastic ($λ=0$), we reject erroneous ~5% of the time (good)
- **Power**: As heteroskedasticity increases ($λ$ grows), we detect it more and more (expected)
- **Sample size dependency**: For the same heteroskedasticity level, larger $N$ makes detection easier

This is perfect for a test statistic (higher power with larger samples), but problematic for a *severity measure* that should give the same reading regardless of $N$.

## 4.2 Formal Test of Parameter Effects (Multilevel ANOVA)

To formally establish which parameters affect $T_{S_{\mathrm{Inf}}}$, we fit a full factorial ANOVA with all experimental factors: sample size (N), heteroskedasticity strength (λ), effect size (β_x), and baseline noise (σ0).

**Note:** With ~300,000 observations, virtually all effects will be statistically significant ($p < 0.001$). We therefore focus on **effect sizes** (η², eta-squared) to assess practical significance:
- η² < 0.01: negligible effect
- η² ∈ [0.01, 0.06): small effect  
- η² ∈ [0.06, 0.14): medium effect
- η² ≥ 0.14: large effect

```{r n-dependence-anova}
# Use RAW simulation data (not aggregated) to preserve within-cell variance
# Convert to factors for ANOVA
hetero_sim_results[, `:=`(
  N_factor = as.factor(N),
  lambda_factor = as.factor(hetero_strength),
  beta_factor = as.factor(beta_x),
  sigma_factor = as.factor(sigma0)
)]

# Fit full factorial ANOVA on raw data
# Note: With ~300k observations, p-values will all be < 0.001
# Focus on eta-squared for practical significance
anova_sinf_full <- aov(T_SInf ~ N_factor * lambda_factor * beta_factor * sigma_factor, 
                       data = hetero_sim_results)
anova_summary_full <- summary(anova_sinf_full)

cat("Full Factorial ANOVA: T_SInf ~ N × λ × β_x × σ0\n")
cat("=" |> rep(70) |> paste(collapse = ""), "\n")
cat(sprintf("Total observations: %d\n\n", nrow(hetero_sim_results)))

# Compute eta-squared for each effect
ss_total <- sum(anova_summary_full[[1]]$`Sum Sq`)
anova_table <- anova_summary_full[[1]]
anova_table$eta_sq <- anova_table$`Sum Sq` / ss_total
anova_table$eta_sq_pct <- round(anova_table$eta_sq * 100, 2)

# Display results with effect sizes
cat("Effect Sizes (η² = Sum Sq / Total Sum Sq):\n")
cat("-" |> rep(70) |> paste(collapse = ""), "\n")
cat(sprintf("%-40s %12s %10s %10s\n", "Effect", "F value", "η²", "η² (%)"))
cat("-" |> rep(70) |> paste(collapse = ""), "\n")

for (i in 1:(nrow(anova_table) - 1)) {  # Exclude Residuals row
  effect_name <- rownames(anova_table)[i]
  f_val <- anova_table$`F value`[i]
  eta_sq <- anova_table$eta_sq[i]
  eta_pct <- anova_table$eta_sq_pct[i]
  
  # Effect size interpretation
  size_interp <- ifelse(eta_sq >= 0.14, "LARGE",
                        ifelse(eta_sq >= 0.06, "Medium",
                               ifelse(eta_sq >= 0.01, "Small", "Negligible")))
  
  cat(sprintf("%-40s %12.1f %10.4f %9.2f%%  [%s]\n", 
              effect_name, f_val, eta_sq, eta_pct, size_interp))
}

# Residuals
cat("-" |> rep(70) |> paste(collapse = ""), "\n")
resid_idx <- nrow(anova_table)
cat(sprintf("%-40s %12s %10.4f %9.2f%%\n", 
            "Residuals", "-", 
            anova_table$eta_sq[resid_idx], 
            anova_table$eta_sq_pct[resid_idx]))
```

```{r anova-interpretation, results='hide'}
# Extract key effects for interpretation (safely handle missing rows)
n_eta <- ifelse("N_factor" %in% rownames(anova_table), 
                anova_table["N_factor", "eta_sq"], NA)
lambda_eta <- ifelse("lambda_factor" %in% rownames(anova_table), 
                     anova_table["lambda_factor", "eta_sq"], NA)
n_lambda_eta <- ifelse("N_factor:lambda_factor" %in% rownames(anova_table), 
                       anova_table["N_factor:lambda_factor", "eta_sq"], NA)

# Check for beta and sigma effects
beta_eta <- ifelse("beta_factor" %in% rownames(anova_table), 
                   anova_table["beta_factor", "eta_sq"], NA)
sigma_eta <- ifelse("sigma_factor" %in% rownames(anova_table), 
                    anova_table["sigma_factor", "eta_sq"], NA)

cat("\n\nKey Findings:\n")
cat("=" |> rep(50) |> paste(collapse = ""), "\n\n")

if (!is.na(n_eta)) {
  cat(sprintf("1. Sample Size (N):           η² = %.4f (%.1f%% of variance)\n", 
              n_eta, n_eta * 100))
}
if (!is.na(lambda_eta)) {
  cat(sprintf("2. Heteroskedasticity (λ):    η² = %.4f (%.1f%% of variance)\n", 
              lambda_eta, lambda_eta * 100))
}
if (!is.na(n_lambda_eta)) {
  cat(sprintf("3. N × λ Interaction:         η² = %.4f (%.1f%% of variance)\n", 
              n_lambda_eta, n_lambda_eta * 100))
}

if (!is.na(beta_eta)) {
  cat(sprintf("4. Effect Size (β_x):         η² = %.4f (%.1f%% of variance)\n", 
              beta_eta, beta_eta * 100))
}
if (!is.na(sigma_eta)) {
  cat(sprintf("5. Baseline Noise (σ0):       η² = %.4f (%.1f%% of variance)\n", 
              sigma_eta, sigma_eta * 100))
}

cat("\n\nInterpretation:\n")
cat("-" |> rep(50) |> paste(collapse = ""), "\n")

if (!is.na(n_eta) && n_eta >= 0.01) {
  cat("→ N explains a meaningful portion of T_SInf variance.\n")
  cat("→ The inferential score is NOT sample-size invariant.\n")
}
if (!is.na(lambda_eta) && lambda_eta >= 0.06) {
  cat("→ λ has a substantial effect (as expected for a heteroskedasticity diagnostic).\n")
}
if (!is.na(n_lambda_eta) && n_lambda_eta >= 0.01) {
  cat("→ The N × λ interaction is meaningful: N-dependence varies with misspecification level.\n")
}
if (!is.na(beta_eta) && beta_eta < 0.01) {
  cat("→ Effect size (β_x) has negligible impact—score is effect-size invariant.\n")
}
if (!is.na(sigma_eta) && sigma_eta < 0.01) {
  cat("→ Baseline noise (σ0) has negligible impact—score is noise-level invariant.\n")
}

cat("\n→ CONCLUSION: T_SInf depends substantially on N, making it unsuitable\n")
cat("   as a direct measure of misspecification SEVERITY across sample sizes.\n")
```

Output: prints the eta-squared breakdown and interpretation for N-dependence.

## 4.3 Visualization of N-Dependence

Having established statistically significant N-dependence, we now visualize how $T_{S_{\mathrm{Inf}}}$ scales with sample size for fixed levels of heteroskedasticity.

```{r n-dependence-sinf-plot}
# Subset to representative parameters (beta_x = 0.5, sigma0 = 1.0)
# Use ALL available lambda values from HETERO_STRENGTH_GRID (0, 0.2, 0.4, ..., 2.0)
subset_params <- hetero_sim_results[beta_x == 0.5 & sigma0 == 1.0]

# Aggregate mean T_SInf
sinf_n_summary <- subset_params[, .(
  mean_sinf = mean(sr_inf),
  mean_T_sinf = mean(T_SInf)
), by = .(N, hetero_strength)]

# Use centralized plotting function from R/60_tables_and_plots.R
print(plot_n_dependence(sinf_n_summary))
```

**Conclusion from Section 4:**
$S_{\mathrm{Inf}}$ is a good *test statistic* (detects heteroskedasticity with high power), but a poor *severity measure* because it scales with $\sqrt{N}$. Two datasets with identical heteroskedasticity can give very different scores just because they have different sample sizes. We need something better.

---

# Section 5: The Solution: Developing an N-Adjusted Reliability Score

Since the inferential score scales too much with $N$, we need a new metric that measures *severity* with reduced sample size dependence. This section develops the **reliability score**—a ratio-based alternative that corrects for finite-sample effects.

The strategy:
1. Start with the **raw SE ratio** (inverse of the inferential score)
2. Demonstrate it has much better N-stability than the inferential score
3. Add a small empirical correction for remaining finite-sample bias
4. Evaluate the degree to which the correction improves N-stability

## 5.0 The Raw Ratio: A Better Starting Point

We flip the inferential score upside down to get the **raw ratio**:
$$R = \frac{\text{SE}_{\text{classic}}}{\text{SE}_{\text{robust}}}$$

Intuition: When HC2 is (correctly) widening the standard error due to heteroskedasticity, $R < 1$. The more heteroskedasticity, the smaller $R$ becomes. This ratio is bounded and interpretable as a severity measure.

```{r compute-ratio-metrics}
# Add ratio metrics post-hoc using the stored SE values
# sr_ratio = se_classic / se_robust (Raw Ratio)
hetero_sim_results[, sr_ratio := se_classic / se_robust]

cat("Raw Ratio (sr_ratio = SE_classic / SE_robust) Summary:\n")
print(summary(hetero_sim_results$sr_ratio))
```

## 5.1 Raw Ratio vs Inferential Score: N-Dependence Comparison

We first compare the N-dependence of the **Raw Ratio** against the **Inferential Score** established in Section 4.

```{r ratio-vs-sinf-comparison, fig.width=12, fig.height=5}
# Use all lambda values for comprehensive comparison
subset_params <- hetero_sim_results[beta_x == 0.5 & sigma0 == 1.0]

# Aggregate both metrics
comparison_summary <- subset_params[, .(
  mean_T_sinf = mean(T_SInf),
  mean_sr_ratio = mean(sr_ratio)
), by = .(N, hetero_strength)]

# Use centralized plotting function from R/60_tables_and_plots.R
plot_tsinf_vs_ratio(comparison_summary)
```

**Key Observation:** The Raw Ratio exhibits dramatically less N-dependence than $T_{S_{\mathrm{Inf}}}$. While $T_{S_{\mathrm{Inf}}}$ shows lines that fan out (diverge) as N increases, the Raw Ratio lines remain approximately parallel across the N range. However, there is still residual finite-sample bias visible as a slight upward drift with N.

## 5.2 Theoretical Basis for Finite-Sample Correction

The residual N-dependence in the Raw Ratio arises from the finite-sample behavior of HC2. Under the null hypothesis (homoskedasticity), the HC2 standard error exhibits finite-sample bias that partially scales with $1/\sqrt{N}$.

The Reliability Score applies a first-order correction to reduce (though not eliminate) this bias, improving cross-sample comparability.

## 5.3 Empirical Estimation of the Scaling Factor

We estimate the optimal scaling constant $c$ by regressing the Raw Ratio on $1/\sqrt{N}$ under the null ($\lambda = 0$).

```{r estimate-scaling-factor}
# Subset to null (lambda = 0) conditions only
null_data <- hetero_sim_results[hetero_strength == 0]

# Compute 1/sqrt(N)
null_data[, inv_sqrt_N := 1 / sqrt(N)]

# Regress raw ratio on 1/sqrt(N)
# sr_ratio = intercept + slope * (1/sqrt(N))
scaling_fit <- lm(sr_ratio ~ inv_sqrt_N, data = null_data)
scaling_summary <- summary(scaling_fit)

cat("Empirical Estimation of Scaling Factor\n")
cat("=" |> rep(60) |> paste(collapse = ""), "\n\n")
cat("Model: sr_ratio = intercept + c * (1/sqrt(N))\n")
cat("Under null (λ = 0), we expect intercept ≈ 1\n\n")

intercept_est <- coef(scaling_fit)[1]
slope_est <- coef(scaling_fit)[2]
c_empirical <- slope_est

cat(sprintf("Intercept:           %.4f (expected: 1.0)\n", intercept_est))
cat(sprintf("Slope (c):           %.4f\n", slope_est))
cat(sprintf("Empirical c:         %.4f\n", c_empirical))
cat(sprintf("\nR²:                  %.4f\n", scaling_summary$r.squared))
```

```{r scaling-visualization, fig.width=8, fig.height=5}
# Aggregate null data by N for cleaner visualization
null_summary <- null_data[, .(
  mean_sr_ratio = mean(sr_ratio),
  se_sr_ratio = sd(sr_ratio) / sqrt(.N)
), by = N]
null_summary[, inv_sqrt_N := 1 / sqrt(N)]

# Use centralized plotting function from R/60_tables_and_plots.R
print(plot_finite_sample_bias(null_summary, intercept_est, slope_est))
```

## 5.4 Sensitivity Analysis: Effect of Model Complexity

We investigate whether the correction factor $c$ is sensitive to model complexity. Since the ANOVA in Section 4.2 showed that $\beta_x$ and $\sigma_0$ have negligible effects on the scores, we compare the finite-sample bias between:

1. **Canonical simple regression**: $Y = \beta_0 + \beta_1 X + \varepsilon$ (p = 2 parameters)
2. **Multiple regression with noise covariate**: $Y = \beta_0 + \beta_1 X + \beta_2 Z + \varepsilon$ (p = 3 parameters)

This sensitivity check determines whether our calibration on simple regression generalizes to more complex models.

```{r sensitivity-analysis}
# Run sensitivity analysis using pipeline functions from R/10_dgp_and_fits.R
# Note: Global seed is set in R/00_config.R; parallel seeds handled via furrr_options(seed = TRUE)
N_test_grid <- c(50, 100, 200, 400, 800, 1600)
n_sims_test <- 10000

cat("Sensitivity Analysis: Effect of Model Complexity on Correction Factor\n")
cat("=" |> rep(60) |> paste(collapse = ""), "\n")
cat(sprintf("Using pipeline functions: run_null_simulation_fast(), run_null_simulation_fast_multiple()\n"))
cat(sprintf("Running %d simulations per N per model type\n\n", n_sims_test))

# Use centralized pipeline function from R/10_dgp_and_fits.R
df_test_results <- run_sensitivity_analysis(
  N_grid = N_test_grid,
  n_sims = n_sims_test,
  hc_type = best_hc,
  verbose = TRUE
)
```

```{r fit-scaling-by-model}
# Fit scaling factor for each model type
simple_fit <- lm(sr_ratio ~ inv_sqrt_N, data = df_test_results[model == "Simple (p=2)"])
multiple_fit <- lm(sr_ratio ~ inv_sqrt_N, data = df_test_results[model == "Multiple (p=3)"])

c_simple <- coef(simple_fit)[2]
c_multiple <- coef(multiple_fit)[2]
intercept_simple <- coef(simple_fit)[1]
intercept_multiple <- coef(multiple_fit)[1]

cat("Sensitivity Check: Effect of Model Complexity\n")
cat("=" |> rep(60) |> paste(collapse = ""), "\n\n")

cat("We compare the finite-sample bias correction factor (c) between:\n")
cat("1. Canonical Simple Regression (p=2)\n")
cat("2. Multiple Regression with Noise Covariate (p=3)\n\n")

cat(sprintf("%-25s %12s %12s\n", "Model", "Empirical c", "R²"))
cat("-" |> rep(50) |> paste(collapse = ""), "\n")
cat(sprintf("%-25s %12.4f %12.4f\n", 
            "Simple (y ~ x, p=2)", 
            c_simple, summary(simple_fit)$r.squared))
cat(sprintf("%-25s %12.4f %12.4f\n", 
            "Multiple (y ~ x + z, p=3)", 
            c_multiple, summary(multiple_fit)$r.squared))
```

```{r sensitivity-plot, fig.width=10, fig.height=5}
# Aggregate by N and model
df_summary <- df_test_results[, .(
  mean_sr_ratio = mean(sr_ratio),
  se_sr_ratio = sd(sr_ratio) / sqrt(.N)
), by = .(N, model, inv_sqrt_N)]

# Use centralized plotting function from R/60_tables_and_plots.R
print(plot_sensitivity_comparison(df_summary, c_simple, c_multiple, 
                                   intercept_simple, intercept_multiple))
```

```{r sensitivity-conclusion}
cat("\nConclusion: Simple Regression Calibration is Conservative\n")
cat("=" |> rep(60) |> paste(collapse = ""), "\n\n")

cat("Adding covariates REDUCES the finite-sample bias:\n")
cat(sprintf("  Simple regression (p=2):   c = %.4f\n", c_simple))
cat(sprintf("  Multiple regression (p=3): c = %.4f\n", c_multiple))
cat(sprintf("  Reduction:                 %.1f%%\n\n", (c_simple - c_multiple) / c_simple * 100))

cat("Implication:\n")
cat("-" |> rep(50) |> paste(collapse = ""), "\n")
cat("Calibrating the Reliability Score on simple regression (c ≈ 0.25)\n")
cat("provides a CONSERVATIVE correction. For models with additional covariates,\n")
cat("the correction slightly over-adjusts, ensuring the metric remains safe\n")
cat("(i.e., it will not underestimate reliability issues).\n\n")

cat(sprintf("We adopt c = %.4f from the simple regression calibration.\n", c_empirical))
cat("\nNote: This provides a conservative finite-sample correction, though complete\n")
cat("N-invariance is not achieved, suggesting higher-order terms may be present.\n")
```

## 5.6 Comparison of Scaling Factors for N-Stability

We now compare the raw ratio against the empirically-adjusted ratio using our calibrated constant derived from the simple regression case:

1. **Raw Ratio** (no correction): $R = \text{SE}_{\text{classic}} / \text{SE}_{\text{robust}}$
2. **Empirical correction** (Reliability Score): $S_{\mathrm{Rel}} = R - c/\sqrt{N}$

```{r compare-scaling-factors}
# Safety Check: Ensure c_empirical is defined from Section 5.3
if (!exists("c_empirical")) stop("c_empirical is missing! Run Section 5.3 first.")

# Compute adjusted ratio using the conservative empirical constant
hetero_sim_results[, sr_ratio_adj := sr_ratio - c_empirical / sqrt(N)]

# Aggregate by N and lambda
scaling_comparison <- hetero_sim_results[, .(
  mean_raw = mean(sr_ratio),
  mean_adjusted = mean(sr_ratio_adj)
), by = .(N, hetero_strength)]

# Compute spread (max - min across N) for each lambda and metric
spread_by_lambda <- scaling_comparison[, .(
  spread_raw = max(mean_raw) - min(mean_raw),
  spread_adjusted = max(mean_adjusted) - min(mean_adjusted)
), by = hetero_strength]

cat("N-Invariance Comparison: Spread (max - min) Across Sample Sizes\n")
cat("=" |> rep(60) |> paste(collapse = ""), "\n\n")
cat(sprintf("Using conservative scaling factor c = %.4f\n\n", c_empirical))
cat(sprintf("%-10s %15s %15s\n", "Lambda", "Raw Ratio", "Adjusted (S_Rel)"))
cat("-" |> rep(60) |> paste(collapse = ""), "\n")

for (i in 1:nrow(spread_by_lambda)) {
  cat(sprintf("%-10.1f %15.4f %15.4f\n",
              spread_by_lambda$hetero_strength[i],
              spread_by_lambda$spread_raw[i],
              spread_by_lambda$spread_adjusted[i]))
}

# Summary statistics
cat("\n")
cat("-" |> rep(60) |> paste(collapse = ""), "\n")
cat(sprintf("%-10s %15.4f %15.4f\n", "Mean",
            mean(spread_by_lambda$spread_raw),
            mean(spread_by_lambda$spread_adjusted)))

# Improvement
improvement <- (mean(spread_by_lambda$spread_raw) - mean(spread_by_lambda$spread_adjusted)) / 
               mean(spread_by_lambda$spread_raw) * 100
cat(sprintf("\nImprovement in mean spread: %.1f%%\n", improvement))
```

```{r scaling-comparison-plot, fig.width=12, fig.height=10}
# Melt for plotting
scaling_long <- melt(scaling_comparison, 
                     id.vars = c("N", "hetero_strength"),
                     measure.vars = c("mean_raw", "mean_adjusted"),
                     variable.name = "Metric", value.name = "Value")

# Convert to data.table if not already
scaling_long <- as.data.table(scaling_long)

scaling_long[, Metric := factor(Metric, 
                                levels = c("mean_raw", "mean_adjusted"),
                                labels = c("Raw Ratio", 
                                          sprintf("Reliability Score (c=%.3f)", c_empirical)))]

# Use centralized plotting function from R/60_tables_and_plots.R
# This creates a 3-panel visualization showing:
# 1. Overlaid comparison (top) - shows flattening effect directly
# 2. Spread reduction by lambda (bottom left) - quantifies improvement
# 3. Slope comparison (bottom right) - shows reduced N-dependence
plot_scaling_comparison(scaling_long)
```

## 5.7 Formal ANOVA Test: Raw Ratio vs Adjusted Scores

We formally test N-dependence for each metric. We expect the N-Effect ($\eta^2$) to be substantially reduced (though not eliminated) for the Reliability Score.

```{r anova-comparison}
# ANOVA on Reliability Score to verify N-invariance
# Note: This is computationally expensive with ~300k observations
cat("Computing ANOVA for Reliability Score N-dependence...\n")

anova_adj <- aov(sr_ratio_adj ~ N_factor + lambda_factor, 
                 data = hetero_sim_results)
ss_adj <- summary(anova_adj)[[1]]
ss_total_adj <- sum(ss_adj$`Sum Sq`)
n_eta_adj <- ifelse("N_factor" %in% rownames(ss_adj), 
                    ss_adj["N_factor", "Sum Sq"] / ss_total_adj, NA)
lambda_eta_adj <- ifelse("lambda_factor" %in% rownames(ss_adj), 
                         ss_adj["lambda_factor", "Sum Sq"] / ss_total_adj, NA)

cat("\nANOVA Effect Size for Reliability Score\n")
cat("=" |> rep(60) |> paste(collapse = ""), "\n\n")

cat(sprintf("%-30s %12s\n", "Factor", "η²"))
cat("-" |> rep(60) |> paste(collapse = ""), "\n")
if (!is.na(n_eta_adj)) {
  cat(sprintf("%-30s %11.4f%%\n", "Sample Size (N)", n_eta_adj * 100))
} else {
  cat(sprintf("%-30s %11s\n", "Sample Size (N)", "[see warning]"))
}
if (!is.na(lambda_eta_adj)) {
  cat(sprintf("%-30s %11.4f%%\n", "Heteroskedasticity (λ)", lambda_eta_adj * 100))
}

cat("\n")
if (!is.na(n_eta_adj) && n_eta_adj < 0.01) {
  cat("✓ The Reliability Score substantially reduces N-dependence (η² < 1%).\n")
  cat(sprintf("  The correction (c = %.3f) provides substantial improvement over raw ratio.\n", c_empirical))
} else if (!is.na(n_eta_adj) && n_eta_adj < 0.06) {
  cat("✓ Reliability Score N-effect is reduced to SMALL (η² < 6%).\n")
  cat(sprintf("  The correction (c = %.3f) reduces but does not eliminate N-dependence.\n", c_empirical))
} else {
  cat("⚠ Warning: N-effect remains noticeable (η² ≥ 6%).\n")
  cat(sprintf("  The correction (c = %.3f) provides partial improvement.\n", c_empirical))
}
```

## 5.8 Classical Coverage Degradation

We visualize how classical CI coverage degrades with increasing heteroskedasticity. This motivates using the Reliability Score to map to coverage loss.

```{r coverage-degradation, fig.width=10, fig.height=5}
# Aggregate coverage by N and hetero_strength
coverage_summary <- hetero_sim_results[, .(
  mean_coverage = mean(coverage) * 100,
  se_coverage = sd(coverage) / sqrt(.N) * 100
), by = .(N, hetero_strength)]

# Use centralized plotting function from R/60_tables_and_plots.R
print(plot_coverage_degradation(coverage_summary))
```

**Key Observation:** Classical coverage degrades systematically with heteroskedasticity. The Reliability Score $S_{\mathrm{Rel}}$ exhibits improved N-stability compared to $S_{\mathrm{Inf}}$, though complete N-invariance is not achieved. This makes it more suitable for comparing heteroskedasticity severity across different sample sizes.

---

# Summary

This notebook completed the validation pipeline using heteroskedastic data. Here's what we found:

## Key Results

1. **S_Inf is a test statistic, not a severity score**
  - $T_{S_{\mathrm{Inf}}}$ increases with sample size even when heteroskedasticity is fixed
  - Eta-squared results show N explains nontrivial variance, so S_Inf is not comparable across datasets

2. **Reliability Score substantially reduces N-dependence**
  $$S_{\mathrm{Rel}} = \frac{\mathrm{SE}_{\text{classic}}}{\mathrm{SE}_{\text{robust}}} - \frac{c}{\sqrt{N}}$$
  - Estimated correction factor: $c \approx 0.23$ (simple regression calibration)
  - The adjustment improves N-stability (47.6% reduction in spread) but does not achieve complete invariance
  - Remaining N-dependence suggests additional finite-sample effects beyond the 1/√N correction

3. **Coverage falls with heteroskedasticity**
  - Classical confidence interval coverage declines as $\lambda$ increases
  - The Reliability Score tracks this deterioration and provides a practical severity metric

## Reproducibility

- All random seeds configured in `R/00_config.R`
- All results saved to `results/` directory
- Run Parts 1–2 in sequence for complete reproducibility
- Complete dataset reproducible by re-running this notebook with identical seed and configuration.
